---
title: "BST 260 - Homework 0"
author: "Alena Figueroa"
format: html
editor: visual
---

## Exercises

(@) Write a Quarto document that defines variables $a = 1, b = -1, c = -2$
and print out the solutions to $f(x) = ax^2 + bx + c = 0$. Do not report complex 
solutions, only real numbers. 

```{r}
#| echo: false
# Define variables a, b, and c
a <- 1
b <- 3
c <- 2

# Compute the discriminate to determine the number of solutions 
discriminate <- b^2 - 4*a*c 

if (discriminate > 0) {
  # Two real-valued solutions
  realsolution1 <- (-b - sqrt(discriminate)) / 2*a
  realsolution2 <- (-b + sqrt(discriminate)) / 2*a
  
  print(strwrap(paste0("There are two real solutions to f(x) = ax\u00B2 + bx + c
                       when a = ", a, ", b = ", b, ", and c = ", c, ".",
                       " The solutions are x = ", realsolution1, " and x = ", 
                       realsolution2, ".")), 
        quote = FALSE)
  
  } else if (discriminate == 0) {
    # One real-valued solution
    realsolution <- -b / 2*a
  
    print(strwrap(paste0("There is one real solution to  f(x) = ax\u00B2 + bx + c 
                       when a = ", a, ", b = ", b, ", and c = ", c, ".",
                       "The solution is x =", realsolution, ".")),
          quote = FALSE)
  
    } else {
      # No real-valued solution
      print(strwrap(paste0("There are no real solutions to f(x) = ax\u00B2 + bx + c 
                       when a = ", a, ", b = ", b, " , and c = ", c, ".")), 
            quote = FALSE)

      }
```


(@) Include a graph of $f(x)$ versus $x$ for $x \in (-5,5)$.

```{r, fig.width = 5, fig.height = 4, fig.align = 'center'}
#| echo: false
x <- seq(-5, 5, length = 300)
plot(x, a*x^2 + b*x + c, type = "l", ylab = "f(x)", ylim = c(-2, 43))
abline(h = 0, lty = 2)
```
The solid black curve in the plot aboves shows the graph of $f(x)$ versus $x$ 
for $x \in (-5,5)$. The points where the curve and horizontal line at $f(x) = 0$
intersect are the real-valued solutions to $f(x) = ax^2 + bx + c = 0$. 


