---
title: "BST 260 - Homework 0"
author: "Alena Figueroa"
format: 
  html:
    self-contained: true
editor: visual
---

## Exercises

(1) Write a Quarto document that defines variables $a = 1, b = -1, c = -2$ and print out the solutions to $f(x) = ax^2 + bx + c = 0$. Do not report complex solutions, only real numbers.

```{r}
#| echo: false
# Define variables a, b, and c
a <- 1
b <- 3
c <- 2

# Compute the discriminate to determine the number of solutions 
discriminate <- b^2 - 4*a*c 

if (discriminate < 0) {
    # No real solution
    print(strwrap(paste0("There are no real solutions to f(x) = ax\u00B2 + bx +
                         c when a = ", a, ", b = ", b, " , and c = ", c, ".")), 
            quote = FALSE)
    solutions <- NA
  } else if (discriminate == 0) {
    # One real solution
    realsolution <- -b / 2*a
  
    print(strwrap(paste0("There is one real solution to  f(x) = ax\u00B2 + bx + 
                         c when a = ", a, ", b = ", b, ", and c = ", c, ".", 
                         "The solution is x =", realsolution, ".")),
          quote = FALSE)
    solutions <- realsolution
    } else {
      # if discriminate > 0 
      # Two real solutions
      realsolution1 <- (-b - sqrt(discriminate)) / 2*a
      realsolution2 <- (-b + sqrt(discriminate)) / 2*a
  
      print(strwrap(paste0("There are two real solutions to f(x) = ax\u00B2 + bx 
                           + c when a = ", a, ", b = ", b, ", and c = ", c, ".", 
                           " The solutions are x = ", realsolution1,
                           " and x = ", realsolution2, ".")), quote = FALSE)
      solutions <- c(realsolution1, realsolution2)
      }
```

(2) Include a graph of $f(x)$ versus $x$ for $x \in (-5,5)$.

```{r, fig.width = 6, fig.height = 5, fig.align = 'center'}
#| echo: false
x <- seq(-5, 5, length = 300)
plot(x, a*x^2 + b*x + c, type = "l", ylab = "f(x)", ylim = c(-2, 43))
abline(h = 0, lty = 2)
points(solutions, rep(0, length(solutions)), pch = 16, col = "cornflowerblue")
```

The solid black curve in the plot above shows the graph of $f(x)$ for $x \in (-5,5)$. The points where the curve and horizontal line at $f(x) = 0$ intersect are the real-valued solutions to $f(x) = ax^2 + bx + c = 0$. That is, the points of intersections provide the real-values of $x$ for which $f(x) = 0$. 
