---
title: "BST 260 - Homework 0"
author: "Alena Figueroa"
format: 
  html:
    self-contained: true
editor: visual
---

## Exercises

(1) Define variables $a$, $b$, and $c$ and print the real-valued solutions
to $f(x) = ax^2 + bx + c = 0$. 

```{r}
#| echo: false

# Define variables a, b, and c
a <- 1
b <- 3
c <- 2

# Compute the discriminate to determine the number of solutions 
discriminate <- b^2 - 4*a*c 

# Condition on whether the discriminate is negative, zero, or positive
# If negative, there is no real solution
# If 0, there is one real solution
# If positive, there are two real solutions 
if (discriminate < 0) {
  
    # No real solution
    # Print a statement regarding the number of real solutions to the function 
    # for the given values of a, b, and c
    print(strwrap(paste0("There are no real solutions to f(x) = ax\u00B2 + bx +
                         c when a = ", a, ", b = ", b, ", and c = ", c, ".")), 
            quote = FALSE)
  
    # Store solution (in this case there are none)
    solutions <- NA
    
  } else if (discriminate == 0) {
    
    # Solve for one real solution 
    realsolution <- -b / (2*a)
    
    # Print solution 
    print(strwrap(paste0("There is one real solution to  f(x) = ax\u00B2 + bx + 
                         c when a = ", a, ", b = ", b, ", and c = ", c, ".", 
                         "The solution is x =", realsolution, ".")),
          quote = FALSE)
    
    # Store solution
    solutions <- realsolution
    
    } else {
      
      # Solve for two real solutions 
      realsolution1 <- (-b - sqrt(discriminate)) / (2*a)
      realsolution2 <- (-b + sqrt(discriminate)) / (2*a)
  
      # Print solutions 
      print(strwrap(paste0("There are two real solutions to f(x) = ax\u00B2 + bx 
                           + c when a = ", a, ", b = ", b, ", and c = ", c, ".", 
                           " The solutions are x = ", realsolution1,
                           " and x = ", realsolution2, ".")), quote = FALSE)
      
      # Store solutions 
      solutions <- c(realsolution1, realsolution2)
      }
```

(2) Plot $f(x)$ versus $x$ for $x \in (-5,5)$ to confirm the solutions found 
in (1).

```{r, fig.width = 6, fig.height = 5, fig.align = 'center'}
#| echo: false

# Store a sequence from -5 to 5 with length 300 in the variable x 
x <- seq(-5, 5, length = 300)

# Plot f(x) = ax^2 + bx + c for each value in the sequence x 
plot(x, a*x^2 + b*x + c, type = "l", ylab = "f(x)", ylim = c(-2, 43))

# Add a horizontal line at f(x) = 0
abline(h = 0, lty = 2)

# Add points where the quadratic function f(x) and horizontal line cross
points(solutions, rep(0, length(solutions)), pch = 16, col = "cornflowerblue")
```

The solid black curve (i.e., the parabola) in the plot above shows the graph of 
the quadratic function $f(x) = ax^2 + bx + c$ for $x \in (-5,5)$ for a given $a$,
$b$, and $c$ (defined above). The points where the curve and horizontal line at 
$f(x) = 0$ intersect are the real-valued solutions to $f(x) = ax^2 + bx + c = 0$.
That is, the points of intersection provide the real-values of $x$ for which the
function $f(x) = ax^2 + bx + c = 0$. If the curve and horizontal line do not 
intersect, there are no real-values of $x$ for which the function
$f(x) = ax^2 + bx + c = 0$. If the curve and horizontal line intersect at one 
point or two points, then there are either one or two real-valued solutions, 
respectively. In the case where $a = 1$, $b = 3$, and $c = 2$, there are two 
real solutions at $x = -2$ and $x = -1$, which is confirmed in the plot since 
there are two points of intersection that correspond to these values of $x$. 
